Index-aware learning of circuits

### 2024-01-23 doc/article/article.tex
revision

\section{MNA and DAEs}
\label{sec:md}
We first look at the system of DAEs that results from MNA when not considering controlled sources. As they are of crucial importance for engineering applications however, we will note whether extensions including controlled sources are available or missing at the appropriate times. Borrowing the notation from Tischendorf\cite{tischendorf1999}, the system of MNA reads
\begin{subequations}
    \label{eq:md_mnai}
    \begin{align}
            \A{C} \ddt \qC \big( \AT{C} \vphi \big) + \A{R} \gR \big( \AT{R} \vphi \big) + \A{L} \i{L} + \A{V} \i{V} + \A{I} \i{s}(t) &= \mb{0}\\ 
            \ddt \phiL \big( \i{L} \big) - \AT{L} \vphi &= \mb{0}\\
            \AT{V} \vphi - \v{s}(t)&= \mb{0},
    \end{align}
\end{subequations}
where $\gR$, $\phiL$ and $\qC$ represent functions modeling resistive, capacitive or inductive devices respectively. The terms for independent current and voltage sources are given by $\i{s}(t)$ and $\v{s}(t)$, while $\vphi$ denotes the vector of nodal potentials and $\i{L} $, $\i{V}$ are the currents flowing through branches containing inductors or voltage sources respectively. The last ingredient is given by the incidence matrices $\A{*}$, where $*$ indicates the device type. These collect the branch to node relations of the underlying electrical network, when considering the branches and nodes as edges and vertices of a directed graph. In the context of design optimization or uncertainty quantification, the device functions $\gR$, $\phiL$ or $\qC$ could e.g.~depend on (subsets of) the parameters $\mb{p}$.

@@

\subsection{Index concepts}
\label{subsec:ic}
\textcolor{red}{Before outlining the dissection index, we want to give a brief introduction to index concepts more generally. There is no unique definition of the index of a DAE, but rather there are multiple related concepts that each possess different strengths and weaknesses\cite{mehrmann2015}. One important aspect that unites these ideas is that they agree in key cases, e.g.~when looking at linear DAEs, and the same also holds true for the dissection index. To emphasize the practical importance of the notion of index, we take a closer look at the perturbation index. It is based on a perturbed version of the DAE \eqref{eq:int_dae} (we leave out the parameters $\mb{p}$ for conciseness)
\begin{align}
    \mb{F}(\mbh{x}', \mbh{x}, t) = \boldsymbol{\varepsilon}(\mbh{x}', \mbh{x}, t), \quad \mbh{x}(0) = \mbh{x}_0, \label{eq:ic_pdae}
\end{align}
where $\varepsilon$ is a sufficiently smooth perturbation, such that the required derivatives exist.
\begin{definition}
    \label{def:ic_pi}
    Let $\mb{x}$ be a solution of the unperturbed DAE, then the DAE is said to have perturbation index $\nu \in \mathbb{N}$, if $\nu$ is the smallest natural number such that, for any sufficiently smooth solution $\mbh{x}$ of \eqref{eq:ic_pdae}, there exists $c \in \mathbb{R}$ with
    \begin{align*}
        \| \mbh{x} - \mb{x} \| \leq c \big( \| \mbh{x}_0 - \mb{x}_0 \| + \| \boldsymbol{\varepsilon} \|_\infty + \| \boldsymbol{\varepsilon}' \|_\infty + \dotsb + \| \boldsymbol{\varepsilon}^{(\nu - 1)} \|_\infty \big)
    \end{align*}
    for an appropriate norm $\| \cdot \|$ and the right hand side small enough.
\end{definition}
The idea behind this definition is to capture the impact of perturbations on the solution, as the name suggests. Usually these pertubations are assumed small, in the sense that $\| \boldsymbol{\varepsilon} \|_\infty \ll 1$, but fast changing, such that $\| \boldsymbol{\varepsilon}^{(\nu)} \|_\infty$ may grow very quickly in $\nu$.}

\paragraph{Example}
\begin{figure}[b]
    \begin{center}
        \includegraphics[width=0.35\textwidth]{vcr}
    \end{center}
    \caption{Small example circuit for illustrating the perturbation index.}
    \label{fig:ic_vcr}
\end{figure}
\textcolor{red}{To illustrate the perturbation index, and also to hint at its relevance for circuit simulation, we consider the small example given in \autoref{fig:ic_vcr}. Applying MNA to the circuit and introducing a perturbation yields the following DAE
\begin{subequations}
    \label{eq:ic_vcr}
    \begin{align}
        C \ddt \hat{\varphi}_1 + \frac{1}{R} \hat{\varphi}_1 + \hat{i}_\mr{V} &= \varepsilon_1(\mbh{x}', \mbh{x}, t) \label{eq:ic_vcra}\\
        -\hat{\varphi}_1 + v_\mr{s}(t) &= \varepsilon_2(\mbh{x}', \mbh{x}, t), \label{eq:ic_vcrb}
    \end{align}
\end{subequations}
where $\mbh{x} = [\hat{\varphi}_1, \hat{i}_\mr{V}]^\T$. Using \eqref{eq:ic_vcrb} we find
\begin{align}
    \hat{\varphi}_1 = v_\mr{s}(t) - \varepsilon_2(\mbh{x}', \mbh{x}, t). \label{eq:ic_vcrphi}
\end{align}
Inserting \eqref{eq:ic_vcrphi} into \eqref{eq:ic_vcra} and rearranging then gives
\begin{align*}
    \hat{i}_\mr{V} &= \varepsilon_1(\mbh{x}', \mbh{x}, t) - C \ddt \big( v_\mr{s}(t) - \varepsilon_2(\mbh{x}', \mbh{x}, t) \big) - \frac{1}{R} \big( v_\mr{s}(t) - \varepsilon_2(\mbh{x}', \mbh{x}, t) \big).
\end{align*}
Noting that the solution $\mb{x}$ to the unperturbed problem follows directly from the perturbed solution by setting $\boldsymbol{\varepsilon} = \mb{0}$, we obtain
\begin{align}
    \| \mbh{x} - \mb{x} \| = \begin{Vmatrix}
        -\varepsilon_2(\mbh{x}', \mbh{x}, t)\\
        \varepsilon_1(\mbh{x}', \mbh{x}, t) + C \ddt \varepsilon_2(\mbh{x}', \mbh{x}, t) + \frac{1}{R} \varepsilon_2(\mbh{x}', \mbh{x}, t)
    \end{Vmatrix}, \label{eq:ic_vcrpi}
\end{align}
so the unperturbed system corresponding to \eqref{eq:ic_vcr} has perturbation index $\nu = 2$ as \eqref{eq:ic_vcrpi} depends on the first derivative of $\varepsilon_2$.}

\paragraph{Index of MNA}
The structure and index of MNA are well understood when only considering independent sources\cite{tischendorf1999} (as in \eqref{eq:md_mnaii}), but also when including controlled sources\cite{estevez2000ijcta}. For the case without controlled sources, there exists the following well-known topological index result.
\begin{theorem}
    \label{th:ic_imna}
    Assuming the matrix-valued functions $\mb{G} \big( \AT{R} \vphi \big)$, $\mb{L}(\i{L})$, $\mb{C} \big( \AT{C} \vphi \big)$ from \eqref{eq:md_dfj} positive definite:
    \begin{enumerate}
        \item The index of \eqref{eq:md_mnaii} is $\nu \leq 2$.
        \item The index of \eqref{eq:md_mnaii} is $\nu \leq 1$, if and only if there are no loops consisting only of capacitors and voltage sources containing at least one voltage source and no cutsets consisting only of inductors and current sources.
    \end{enumerate}
\end{theorem}
The result can be found in terms of the tractability index\cite{tischendorf1999}, the differentiation index\cite{estevez2000ijcta} and the dissection index\cite{jansen2014}. We again note that there also exist extensive results about when the index of MNA including controlled sources does not exceed $\nu = 2$\cite{estevez2000ijcta}.

\section{Dissection index}
\label{sec:di}
\textcolor{red}{We now focus on the dissection index. As mentioned in \autoref{sec:int}, the key property of this index concept is that it enables the decoupling of a DAE into an ODE and a set of purely algebraic equations. This is clearly different from the perturbation index, however other index concepts such as the tractability and differentiation indices also make such a decoupling possible. Still, the dissection index maintains some advantages over these concepts, as it provides a simple algorithmic procedure for the decoupling that is similar to the tractability index, and it poses less strict smoothness assumptions. In the case of MNA without controlled sources \eqref{eq:md_mnaii}, it is even possible to find a purely topological decoupling based on the dissection index\cite{jansen2014}. We will not make use of this topological decoupling here however, but rather consider the dissection index for a more general class of DAEs, to formulate the assumptions that are necessary for our method to work also for DAEs other than \eqref{eq:md_mnaii}.}

@@

\subsection{Index two case}
\label{subsec:itc}
As the example from \autoref{fig:ic_vcr} illustrates, there are many DAEs, including those described by MNA, which can have an index higher than one. In these cases the dissection index proceeds by introducing additional basis functions and continuing the splitting process in a similar fashion. We begin by focusing on the algebraic equation \eqref{eq:ioc_iob}, and consider basis functions $\mbb{P}(\mb{x})$, $\mbb{Q}(\mb{x})$, $\mbb{V}(\mb{x})$, $\mbb{W}(\mb{x})$ of $\mbb{K}_\mb{Q}(\mb{x})$, defined analogous to the ones for $\mb{M}(\mb{x})$. This allows us to further split the algebraic variables $\mbb{x}$ as follows
\begin{align}
    \mbb{x} = \mbb{P}(\mb{x}) \mbb{x}_\mb{P} + \mbb{Q}(\mb{x}) \mbb{x}_\mb{Q}. \label{eq:itc_xb}
\end{align}
Inserting this splitting into \eqref{eq:ioc_iob}, and multiplying once by $\mbb{V}^\T (\mb{x})$ and $\mbb{W}^\T (\mb{x})$ each from the left, splits the algebraic equation into two parts
\begin{subequations}
    \label{eq:itc_iti}
    \begin{align}
        \mbb{V}^\T (\mb{x}) \mbb{K}_\mb{P}(\mb{x}) \mbt{x} + \mbb{V}^\T (\mb{x}) \mbb{K}_\mb{Q}(\mb{x}) \mbb{P}(\mb{x}) \mbb{x}_\mb{P} + \mbb{V}^\T (\mb{x}) \mbb{f}(t) &= \mb{0} \label{eq:itc_itia}\\
        \mbb{W}^\T (\mb{x}) \mbb{K}_\mb{P}(\mb{x}) \mbt{x} + \mbb{W}^\T (\mb{x}) \mbb{f}(t) &= \mb{0}. \label{eq:itc_itib}
    \end{align}
\end{subequations}
We can now also split the differential variables $\mbt{x}$ further, by using basis functions $\mbt{P}(\mb{x})$ and $\mbt{Q}(\mb{x})$ of $\mbb{W}^\T (\mb{x}) \mbb{K}_\mb{P}(\mb{x})$
\begin{align}
    \mbt{x} = \mbt{P}(\mb{x}) \mbt{x}_\mb{P} +
    \mbt{Q}(\mb{x}) \mbt{x}_\mb{Q}, \label{eq:itc_xt}
\end{align}
and inserting this splitting into the second algebraic equation \eqref{eq:itc_itib} yields
\begin{align}
    \mbb{W}^\T (\mb{x}) \mbb{K}_\mb{P}(\mb{x}) \mbt{P}(\mb{x}) \mbt{x}_\mb{P} + \mbb{W}^\T (\mb{x}) \mbb{f}(t) &= \mb{0}. \label{eq:itc_xtp}
\end{align}
We now make the following two assumptions.
\textcolor{red}{\begin{assumption}
    \label{as:itc_xtpdi}
    The matrix product $\mbb{W}^\T (\mb{x}) \mbb{K}_\mb{P}(\mb{x}) \mbt{P}(\mb{x}) \mbt{x}_\mb{P}$ is regular.
\end{assumption}
\begin{assumption}
    \label{as:itc_xtp}
    The basis functions $\mbt{P}$ and $\mbt{Q}$ of $\mbb{W}^\T (\mb{x}) \mbb{K}_\mb{P}(\mb{x})$ are constant and \eqref{eq:itc_xtp} possesses a locally unique solution for $\mbt{x}_\mb{P}$ in terms of $\mbt{x}_\mb{Q}$ and $t$.
\end{assumption}
We note that while \autoref{as:itc_xtpdi} is equivalent to the DAE not being underdetermined\cite{jansen2014}, \autoref{as:itc_xtp} is more practical and important for the implementation, but not for the dissection index itself. In fact our approach still works if the basis functions $\mbt{P}(\mbt{x}_\mb{Q})$ and $\mbt{Q}(\mbt{x}_\mb{Q})$ depend on $\mbt{x}_\mb{Q}$, however we focus on the stronger assumption here, since it shortens the expressions in the following without impacting the general idea and \autoref{re:dim_xtp} shows that MNA fulfills an even stronger condition than \autoref{as:itc_xtp}.}

@@

There is yet another, maybe less expected, benefit that might occur. Looking at the decoupled system from \eqref{eq:itc_do2} we find that the resistance parameter $R$ and the inductance parameter $L$ only appear in the algebraic equation. In terms of our original goal of speeding up design optimization or uncertainty quantification, where a lot of solutions for varying parameter values are required, this means the following: instead of having to solve the full system for a given combination of $R$ and $L$, we can instead simply solve the algebraic equation to obtain the full solution. While the algebraic equation might be more complicated than the simple linear relations of \eqref{eq:itc_do1} and \eqref{eq:itc_do2} in general, solving it is almost certainly much faster than having to integrate the entire system in time. This becomes an even bigger advantage when knowledge about the solution is only required for specific points in time, since the algebraic equation may be solved pointwise. \textcolor{red}{As of now however, we have no knowledge of an easy way to automatically determine which parameters appear in the ODE. But when combined with a sequential learning strategy, such as the one outlined in \autoref{sec:ne}, there may still be computational savings due to the learning method requiring less samples for the parameters not appearing in the ODE.}

@@

\paragraph{Rectifier circuit}
\begin{figure}[t]
    \begin{center}
        \includegraphics[width=0.5\textwidth]{fwr2}
    \end{center}
    \caption{Full wave rectifier circuit.}
    \label{fig:ne_fwr2}
\end{figure}

>>>>

\section{MNA and DAEs}
\label{sec:md}
We first look at the system of DAEs that results from MNA when not considering controlled sources. As they are of crucial importance for engineering applications however, we will note whether extensions including controlled sources are available or missing at the appropriate times. Borrowing the notation from Tischendorf\cite{tischendorf1999}, the system of MNA reads
\begin{subequations}
    \label{eq:md_mnai}
    \begin{align}
            \A{C} \ddt \qC \big( \AT{C} \vphi \big) + \A{R} \gR \big( \AT{R} \vphi \big) + \A{L} \i{L} + \A{V} \i{V} + \A{I} \i{s}(t) &= \mb{0}\\ 
            \ddt \phiL \big( \i{L} \big) - \AT{L} \vphi &= \mb{0}\\
            \AT{V} \vphi - \v{s}(t)&= \mb{0},
    \end{align}
\end{subequations}
where the left hand side as a whole corresponds to $\mb{F}$ in \eqref{eq:int_dae}, and the solution variables are given by $\mb{x} = [\vphi, \i{L}, \i{V}]^\T$. The functions $\gR$, $\phiL$ and $\qC$ model resistive, capacitive or inductive devices respectively, that may each depend on the parameters $\mb{p}$. The terms for independent current and voltage sources are given by $\i{s}(t)$ and $\v{s}(t)$, while $\vphi$ denotes the vector of nodal potentials and $\i{L} $, $\i{V}$ are the currents flowing through branches containing inductors or voltage sources respectively. The last ingredient is given by the incidence matrices $\A{*}$, where $*$ indicates the device type. These collect the branch to node relations of the underlying electrical network, when considering the branches and nodes as edges and vertices of a directed graph.

@@

\subsection{Index concepts}
\label{subsec:ic}
\textcolor{red}{Before outlining the dissection index, we want to give a brief introduction to index concepts more generally. There are multiple definitions of the index of a DAE, along with related index concepts that each possess different strengths and weaknesses\cite{mehrmann2015}. One important aspect that unites these ideas is that they agree in key cases, e.g.~when looking at linear DAEs, and the same also holds true for the dissection index. To emphasize the practical importance of the notion of index, we take a closer look at the perturbation index. It is based on a perturbed version of the DAE \eqref{eq:int_dae} (we leave out the parameters $\mb{p}$ for conciseness)
\begin{align}
    \mb{F}(\mbh{x}', \mbh{x}, t) = \boldsymbol{\varepsilon}(t), \quad \mbh{x}(0) = \mbh{x}_0, \label{eq:ic_pdae}
\end{align}
where $\varepsilon$ is a sufficiently smooth perturbation, such that the required derivatives exist.
\begin{definition}
    \label{def:ic_pi}
    Let $\mb{x}$ be a solution of the unperturbed DAE, then the DAE is said to have perturbation index $\nu \in \mathbb{N}$, if $\nu$ is the smallest natural number such that, for any sufficiently smooth solution $\mbh{x}$ of \eqref{eq:ic_pdae}, there exists $c \in \mathbb{R}$ with
    \begin{align*}
        \| \mbh{x} - \mb{x} \| \leq c \big( \| \mbh{x}_0 - \mb{x}_0 \| + \| \boldsymbol{\varepsilon} \|_\infty + \| \boldsymbol{\varepsilon}' \|_\infty + \dotsb + \| \boldsymbol{\varepsilon}^{(\nu - 1)} \|_\infty \big)
    \end{align*}
    for an appropriate norm $\| \cdot \|$ and the right hand side small enough.
\end{definition}
The idea behind this definition is to capture the impact of perturbations on the solution, as the name suggests. Usually these pertubations are assumed small, in the sense that $\| \boldsymbol{\varepsilon} \|_\infty \ll 1$, but fast changing, such that $\| \boldsymbol{\varepsilon}^{(\nu)} \|_\infty$ may grow very quickly in $\nu$.}

\paragraph{Example}
\begin{figure}[b]
    \begin{center}
        \includegraphics[width=0.35\textwidth]{vcr}
    \end{center}
    \caption{Small example circuit for illustrating the perturbation index.}
    \label{fig:ic_vcr}
\end{figure}
\textcolor{red}{To illustrate the perturbation index, and also to hint at its relevance for circuit simulation, we consider the small example given in \autoref{fig:ic_vcr}. Applying MNA to the circuit and introducing a perturbation yields the following DAE
\begin{subequations}
    \label{eq:ic_vcr}
    \begin{align}
        C \ddt \hat{\varphi}_1 + \frac{1}{R} \hat{\varphi}_1 + \hat{i}_\mr{V} &= \varepsilon_1(t) \label{eq:ic_vcra}\\
        -\hat{\varphi}_1 + v_\mr{s}(t) &= \varepsilon_2(t), \label{eq:ic_vcrb}
    \end{align}
\end{subequations}
where $\mbh{x} = [\hat{\varphi}_1, \hat{i}_\mr{V}]^\T$. Using \eqref{eq:ic_vcrb} we find
\begin{align}
    \hat{\varphi}_1 = v_\mr{s}(t) - \varepsilon_2(t). \label{eq:ic_vcrphi}
\end{align}
Inserting \eqref{eq:ic_vcrphi} into \eqref{eq:ic_vcra} and rearranging then gives
\begin{align*}
    \hat{i}_\mr{V} &= \varepsilon_1(t) - C \ddt \big( v_\mr{s}(t) - \varepsilon_2(t) \big) - \frac{1}{R} \big( v_\mr{s}(t) - \varepsilon_2(t) \big).
\end{align*}
Noting that the solution $\mb{x}$ to the unperturbed problem follows directly from the perturbed solution by setting $\boldsymbol{\varepsilon} = \mb{0}$, we obtain
\begin{align}
    \| \mbh{x} - \mb{x} \| = \begin{Vmatrix}
        -\varepsilon_2(t)\\
        \varepsilon_1(t) + C \ddt \varepsilon_2(t) + \frac{1}{R} \varepsilon_2(t)
    \end{Vmatrix}, \label{eq:ic_vcrpi}
\end{align}
so the unperturbed system corresponding to \eqref{eq:ic_vcr} has perturbation index $\nu = 2$ as \eqref{eq:ic_vcrpi} depends on the first derivative of $\varepsilon_2$.}

\paragraph{Index of MNA}
The structure and index of MNA are well understood when only considering independent sources\cite{tischendorf1999} (as in \eqref{eq:md_mnaii}), but also when including controlled sources\cite{estevez2000ijcta}. For the case without controlled sources, there exists the following well-known topological index result.
\begin{theorem}
    \label{th:ic_imna}
    Assuming the matrix-valued functions $\mb{G} \big( \AT{R} \vphi \big)$, $\mb{L}(\i{L})$, $\mb{C} \big( \AT{C} \vphi \big)$ from \eqref{eq:md_dfj} positive definite:
    \begin{enumerate}
        \item The index of \eqref{eq:md_mnaii} is $\nu \leq 2$.
        \item The index of \eqref{eq:md_mnaii} is $\nu \leq 1$, if and only if there are no loops consisting only of capacitors and voltage sources containing at least one voltage source and no cutsets consisting only of inductors and current sources.
    \end{enumerate}
\end{theorem}
The result can be found in terms of the perturbation index\cite{gunther1996}, the tractability index\cite{tischendorf1999}, the differentiation index\cite{estevez2000ijcta} and the dissection index\cite{jansen2014}. We again note that there also exist extensive results about when the index of MNA including controlled sources does not exceed $\nu = 2$\cite{estevez2000ijcta}.

\section{Dissection index}
\label{sec:di}
\textcolor{red}{We focus on the dissection index, as it enables the decoupling of a DAE into an ODE and a set of purely algebraic equations. This is conceptually different from the perturbation index, however other index concepts such as the tractability and differentiation indices also use decoupling strategies. Still, the dissection index maintains some advantages over these concepts, as it provides a simple algorithmic procedure for the decoupling that is similar to the tractability index, but poses less strict smoothness assumptions. In the case of MNA without controlled sources \eqref{eq:md_mnaii}, it is even possible to find a purely topological decoupling based on the dissection index\cite{jansen2014}. We will not make use of this topological decoupling in the derivation however, but rather consider the dissection index for a more general class of DAEs, to formulate the assumptions that are necessary for our method to work also for DAEs other than \eqref{eq:md_mnaii}.}

@@

\subsection{Index two case}
\label{subsec:itc}
As the example from \autoref{fig:ic_vcr} illustrates, there are many DAEs, including those described by MNA, which can have an index higher than one. In these cases the dissection index proceeds by introducing additional basis functions and continuing the splitting process in a similar fashion. We begin by focusing on the algebraic equation \eqref{eq:ioc_iob}, and consider basis functions $\mbb{P}(\mb{x})$, $\mbb{Q}(\mb{x})$, $\mbb{V}(\mb{x})$, $\mbb{W}(\mb{x})$ of $\mbb{K}_\mb{Q}(\mb{x})$, defined analogous to the ones for $\mb{M}(\mb{x})$. This allows us to further split the algebraic variables $\mbb{x}$ as follows
\begin{align}
    \mbb{x} = \mbb{P}(\mb{x}) \mbb{x}_\mb{P} + \mbb{Q}(\mb{x}) \mbb{x}_\mb{Q}. \label{eq:itc_xb}
\end{align}
Inserting this splitting into \eqref{eq:ioc_iob}, and multiplying once by $\mbb{V}^\T (\mb{x})$ and $\mbb{W}^\T (\mb{x})$ each from the left, splits the algebraic equation into two parts
\begin{subequations}
    \label{eq:itc_iti}
    \begin{align}
        \mbb{V}^\T (\mb{x}) \mbb{K}_\mb{P}(\mb{x}) \mbt{x} + \mbb{V}^\T (\mb{x}) \mbb{K}_\mb{Q}(\mb{x}) \mbb{P}(\mb{x}) \mbb{x}_\mb{P} + \mbb{V}^\T (\mb{x}) \mbb{f}(t) &= \mb{0} \label{eq:itc_itia}\\
        \mbb{W}^\T (\mb{x}) \mbb{K}_\mb{P}(\mb{x}) \mbt{x} + \mbb{W}^\T (\mb{x}) \mbb{f}(t) &= \mb{0}. \label{eq:itc_itib}
    \end{align}
\end{subequations}
We can now also split the differential variables $\mbt{x}$ further, by using basis functions $\mbt{P}(\mb{x})$ and $\mbt{Q}(\mb{x})$ of $\mbb{W}^\T (\mb{x}) \mbb{K}_\mb{P}(\mb{x})$
\begin{align}
    \mbt{x} = \mbt{P}(\mb{x}) \mbt{x}_\mb{P} +
    \mbt{Q}(\mb{x}) \mbt{x}_\mb{Q}, \label{eq:itc_xt}
\end{align}
and inserting this splitting into the second algebraic equation \eqref{eq:itc_itib} yields
\begin{align}
    \mbb{W}^\T (\mb{x}) \mbb{K}_\mb{P}(\mb{x}) \mbt{P}(\mb{x}) \mbt{x}_\mb{P} + \mbb{W}^\T (\mb{x}) \mbb{f}(t) &= \mb{0}. \label{eq:itc_xtp}
\end{align}
We now make the following two assumptions.
\textcolor{red}{\begin{assumption}
    \label{as:itc_xtpdi}
    The matrix product $\mbb{W}^\T (\mb{x}) \mbb{K}_\mb{P}(\mb{x}) \mbt{P}(\mb{x}) \mbt{x}_\mb{P}$ is regular.
\end{assumption}
\begin{assumption}
    \label{as:itc_xtp}
    The basis functions $\mbt{P}$ and $\mbt{Q}$ of $\mbb{W}^\T (\mb{x}) \mbb{K}_\mb{P}(\mb{x})$ are constant and \eqref{eq:itc_xtp} possesses a locally unique solution for $\mbt{x}_\mb{P}$ in terms of $\mbt{x}_\mb{Q}$ and $t$.
\end{assumption}
We note that while \autoref{as:itc_xtpdi} is equivalent to the DAE not being underdetermined\cite{jansen2014}, \autoref{as:itc_xtp} is rather important for the implementation, but not for the dissection index itself. In fact our approach still works if the basis functions $\mbt{P}(\mbt{x}_\mb{Q})$ and $\mbt{Q}(\mbt{x}_\mb{Q})$ depend on $\mbt{x}_\mb{Q}$, however we focus on the stronger assumption here, since it shortens the expressions in the following without impacting the general idea and \autoref{re:dim_xtp} shows that MNA fulfills an even stronger condition than \autoref{as:itc_xtp}.}

@@

There is yet another, maybe less expected, benefit that might occur. Looking at the decoupled system from \eqref{eq:itc_do2} we find that the resistance parameter $R$ and the inductance parameter $L$ only appear in the algebraic equation. In terms of our original goal of speeding up design optimization or uncertainty quantification, where a lot of solutions for varying parameter values are required, this means the following: instead of having to solve the full system for a given combination of $R$ and $L$, we can instead simply solve the algebraic equation to obtain the full solution. While the algebraic equation might be more complicated than the simple linear relations of \eqref{eq:itc_do1} and \eqref{eq:itc_do2} in general, solving it is almost certainly much faster than having to integrate the entire system in time. This becomes an even bigger advantage when knowledge about the solution is only required for specific points in time, since the algebraic equation may be solved pointwise. \textcolor{red}{As of now, we have no easy way to automatically determine which parameters appear in the ODE. But when combined with a sequential learning strategy, such as the one outlined in \autoref{sec:ne}, there may still be computational savings due to the learning method requiring less samples for the parameters not appearing in the ODE.}

@@

\paragraph{Rectifier circuit}
\begin{figure}[t]
    \begin{center}
        \includegraphics[width=0.5\textwidth]{fwr2}
    \end{center}
    \caption{Full wave rectifier circuit. The resistances of the diodes depend on the parameters $\mb{p} = [T_1, T_2]^\T$.}
    \label{fig:ne_fwr2}
\end{figure}

### 2024-01-23 doc/article/letter.tex
revision

        \item[3)] \textcolor{TUDa-0c}{Page 8, line 34-35: Can we always decouple the parameters into those that (may) appear in the algebraic equations and those that (may) appear in the ODE? Are there topological criteria for this? And is your approach reasonable if such a classification does not exist?}
        \begin{itemize}
            \item For the topological basis functions mentioned in section 3 this might be possible, but we have not explored this in more detail.
            \item We pointed this out more explicitly in the manuscript and remarked that for sequential learning approaches, such as the one used in the paper, one can still benefit from parameters not appearing in the ODE without any explicit knowledge.\\
            ``\textcolor{red}{As of now however, we have no knowledge of an easy way to automatically determine which parameters appear in the ODE. But when combined with a sequential learning strategy, such as the one outlined in \dots, there may still be computational savings due to the learning method requiring less samples for the parameters not appearing in the ODE.}'' (p.9)
            \item The approach is reasonable independently of this, since the two main benefits (a reduced number of variables to learn and better accuracy in the algebraic constraints) are independent of this anyhow.
        \end{itemize}

@@

        \item[6)] \textcolor{TUDa-0c}{You emphasize that the approach is independent of the machine learning method used. Is it also applicable for any other decoupling of DAEs into an inherent ODE and the algebraic constraints, also if it is based on other index concepts?}
        \begin{itemize}
            \item In principle it is possible to use any index concept that provides such a decoupling. There are however clear advantages when using the dissection index, in particular in the context of MNA, see also the answer to 5).
            \item We added a remark about this at the beginning of section 3 and also motivated the dissection index more strongly.\\
            ``\textcolor{red}{We now focus on the dissection index. As mentioned in \dots, the key property of this index concept is that it enables the decoupling of a DAE into an ODE and a set of purely algebraic equations. This is clearly different from the perturbation index, however other index concepts such as the tractability and differentiation indices also make such a decoupling possible. Still, the dissection index maintains some advantages over these concepts, as it provides a simple algorithmic procedure for the decoupling that is similar to the tractability index, and it poses less strict smoothness assumptions. In the case of MNA without controlled sources \dots, it is even possible to find a purely topological decoupling based on the dissection index. We will not make use of this topological decoupling here however, but rather consider the dissection index for a more general class of DAEs, to formulate the assumptions that are necessary for our method to work also for DAEs other than \dots.}'' (p.4)
        \end{itemize}
    \end{itemize}

@@

            \item Lastly, we removed the alternative approach (previous section 5.1) to achieve a clearer presentation of our main approach.
        \end{itemize}

>>>>

        \item[3)] \textcolor{TUDa-0c}{Page 8, line 34-35: Can we always decouple the parameters into those that (may) appear in the algebraic equations and those that (may) appear in the ODE? Are there topological criteria for this? And is your approach reasonable if such a classification does not exist?}
        \begin{itemize}
            \item For the topological basis functions mentioned in section 3 this might be possible, but we have not explored this in more detail.
            \item We pointed this out more explicitly in the manuscript and remarked that for sequential learning approaches, such as the one used in the paper, one can still benefit from parameters not appearing in the ODE without any explicit knowledge.\\
            ``\textcolor{red}{As of now however, we have no knowledge of an easy way to automatically determine which parameters appear in the ODE. But when combined with a sequential learning strategy, such as the one outlined in \dots, there may still be computational savings due to the learning method requiring less samples for the parameters not appearing in the ODE.}'' (p.9)
            \item The approach is also reasonable if the number of parameters cannot be reduced by the decoupling, since two significant advantages remain: better accuracy in the algebraic constraints and a reduced number of (differential) variables to learn.
        \end{itemize}

@@

        \item[6)] \textcolor{TUDa-0c}{You emphasize that the approach is independent of the machine learning method used. Is it also applicable for any other decoupling of DAEs into an inherent ODE and the algebraic constraints, also if it is based on other index concepts?}
        \begin{itemize}
            \item In principle it is possible to use any index concept that provides such a decoupling. There are however clear advantages when using the dissection index, in particular in the context of MNA, see also the answer to 5).
            \item We added a remark about this at the beginning of section 3 and also motivated the dissection index more strongly.\\
            ``\textcolor{red}{We focus on the dissection index, as it enables the decoupling of a DAE into an ODE and a set of purely algebraic equations. This is conceptually different from the perturbation index, however other index concepts such as the tractability and differentiation indices also use decoupling strategies. Still, the dissection index maintains some advantages over these concepts, as it provides a simple algorithmic procedure for the decoupling that is similar to the tractability index, but poses less strict smoothness assumptions. In the case of MNA without controlled sources \dots, it is even possible to find a purely topological decoupling based on the dissection index. We will not make use of this topological decoupling in the derivation however, but rather consider the dissection index for a more general class of DAEs, to formulate the assumptions that are necessary for our method to work also for DAEs other than \dots.}'' (p.4)
        \end{itemize}
    \end{itemize}

@@

            \item Lastly, we removed the alternative approach (previous section 5.1) to achieve a clearer presentation of our main message.
        \end{itemize}
